\subsection{🚀 Release v0.9.2 ``Unified Urubu''
Highlights}\label{release-v0.9.2-unified-urubu-highlights}

\begin{itemize}
\tightlist
\item
  \textbf{Unified All-in-One Build:} The \texttt{-\/-all} flag now runs
  all major build steps in strict order: DOCX (and Markdown) → PDF →
  Jupyter Book HTML → Custom HTML web output. All output formats are
  copied to
  \texttt{docs/sources/\textless{}notebook\_stem\textgreater{}/} for
  each notebook.
\item
  \textbf{Jupyter Book HTML Improvements:} Built into a temp directory,
  then copied to both \texttt{docs/jupyter/} and
  \texttt{docs/sources/jupyterbook\_html/} for easy access and
  deployment.
\item
  \textbf{Documentation Overhaul:} \texttt{README.md} and
  \texttt{build.md} have been fully rewritten to reflect the new build
  workflow, directory structure, and usage patterns.
\item
  \textbf{Output Directory Clarity:} All outputs are organized in
  \texttt{\_build/} (intermediate) and \texttt{docs/} (final,
  published), with clear subfolders for each format. Downloadable
  sources for each notebook are always available in
  \texttt{docs/sources/\textless{}notebook\_stem\textgreater{}/}.
\item
  \textbf{Robust Image Handling:} All referenced images are collected
  into \texttt{\_build/images/} and linked consistently in all output
  formats.
\item
  \textbf{No Automatic Cleanup:} The build system no longer deletes any
  files or folders automatically; manual cleanup is recommended if
  needed.
\end{itemize}

See the \texttt{releases/} folder for full historical release notes.

For more details and instructions, see the
\href{https://github.com/dannycab/modern-classical-mechanics/releases}{Release
Notes}.
